R package glmm

### 2014-05-28 CompareExamples/glmmexamples.tex
Added documentation and updated murder data
Added to glmm examples document

First, I fit the data using glmer in lme4. Then I fit the data using glmm with $m=10^4$. This took 16.75 minutes on my netbook. The point estimates and the standard errors were very similar between the two  methods of model-fitting. (The data did need to be reformatted, which took only a couple minutes)

@@

\section{Salamander}

@@

Here are the results with $m=10^4$ but I forgot to set the seed.
\begin{verbatim}
> summary(sal.glmm2)

@@

>>>>

First, I fit the data using glmer in lme4. Then I fit the data using glmm with $m=10^4$. This took 16.75 minutes on my netbook. The point estimates and the standard errors were very similar between the two  methods of model-fitting. The documentation says that the default method of model-fitting uses the Laplace approximation, but Charlie wonders if they actually use numerical integration for simple problems like this.

@@

\section{Salamander}
I fit the salamander data set using  my glmm package and compared those results to the results of Yun Ju Sung and Charlie Geyer.

@@

Here are the results with $m=10^4$ but I forgot to set the seed. At least it helps us see a bit of the variability.
\begin{verbatim}
> summary(sal.glmm2)

@@

\section{Murder}
Subjects were asked how many victims of homicide they personally knew. The data set is from Agresti's \textit{Categorical Data Analysis} book. I found similar answers from my glmm package as from glmmPQL.

The results from my package:

\begin{verbatim}
set.seed(1234)
murder.glmm<- glmm(y~race ,random=list(~0+black,~0+white), varcomps.names=c("black","white"), data=murder ,family.glmm=poisson.glmm,m=10^4)

> summary(murder.glmm)

Call:
glmm(fixed = y ~ race, random = list(~0 + black, ~0 + white), 
    varcomps.names = c("black", "white"), data = murder, family.glmm = poisson.glmm, 
    m = 10^4)

Fixed Effects:
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)  0.42003    0.06428   6.534 6.40e-11 ***
raceWhite   -0.33179    0.07021  -4.726 2.29e-06 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Variance Components for Random Effects (P-values are one-tailed):
       Estimate Std. Error z value Pr(>|z|)/2
black 6.468e-10  3.452e-09   0.187      0.426
white 2.667e-10  1.596e-09   0.167      0.434
\end{verbatim}

When I make the sample size drastically smaller so that the analysis takes only about a minute, the results are very similar still.
\begin{verbatim}
> summary(murder.glmm)

Call:
glmm(fixed = y ~ race, random = list(~0 + black, ~0 + white), 
    varcomps.names = c("black", "white"), data = murder, family.glmm = poisson.glmm, 
    m = 10^2)

Fixed Effects:
            Estimate Std. Error z value Pr(>|z|)    
(Intercept)  0.42003    0.06428   6.534 6.39e-11 ***
raceWhite   -0.33179    0.07021  -4.726 2.29e-06 ***
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

Variance Components for Random Effects (P-values are one-tailed):
       Estimate Std. Error z value Pr(>|z|)/2
black 1.190e-09  1.718e-09   0.693      0.244
white 1.070e-08  1.440e-08   0.743      0.229

The murder model resulting from glmmPQL:
\begin{verbatim}
> murder.pql<-glmmPQL(y~race,random=~1|race,data=murder,family=poisson)iteration 1
> summary(murder.pql)
Linear mixed-effects model fit by maximum likelihood
 Data: murder 
  AIC BIC logLik
   NA  NA     NA

Random effects:
 Formula: ~1 | race
         (Intercept)  Residual
StdDev: 9.157794e-06 0.4647694

Variance function:
 Structure: fixed weights
 Formula: ~invwt 
Fixed effects: y ~ race 
                 Value  Std.Error   DF   t-value p-value
(Intercept)  0.4200335 0.02989812 1306  14.04883       0
raceWhite   -0.3317900 0.03265396    0 -10.16079     NaN
 Correlation: 
          (Intr)
raceWhite -0.916

Standardized Within-Group Residuals:
       Min         Q1        Med         Q3        Max 
-0.9104427 -0.1899269 -0.1899269 -0.1899269 12.1624898

Number of Observations: 1308
Number of Groups: 2 
Warning message:
In pt(-abs(tTable[, "t-value"]), tTable[, "DF"]) : NaNs produced

%\section{Booth and Hobert}
%Here are the results from my glmm package fitting a model to the Booth and Hobert data (with $m=10^4$).
%\begin{verbatim}
%> set.seed(1234)
%> mod.mcml1<-glmm(y~0+x1,list(y~0+z1),varcomps.names=c("z1"),data=BoothHobert,family.glmm=bernoulli.glmm,m=10^4,doPQL=TRUE)
%[1] "PQL estimates have been identified."
%> summary(mod.mcml1)
%
%Call:
%glmm(fixed = y ~ 0 + x1, random = list(y ~ 0 + z1), varcomps.names = c("z1"), 
%    data = BoothHobert, family.glmm = bernoulli.glmm, m = 10^4, 
%    doPQL = TRUE)
%
%Fixed Effects:
%   Estimate Std. Error z value Pr(>|z|)    
%x1    6.225      1.422   4.377  1.2e-05 ***
%---
%Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
%
%
%Variance Components for Random Effects (P-values are one-tailed):
%   Estimate Std. Error z value Pr(>|z|)/2
%z1    1.907      1.814   1.051      0.147
%
%
%\end{verbatim}
%
%Here are the results from glmmPQL:
%\begin{verbatim}
%> summary(bh)
%Linear mixed-effects model fit by maximum likelihood
% Data: BoothHobert 
%  AIC BIC logLik
%   NA  NA     NA
%
%Random effects:
% Formula: ~1 | z1
%        (Intercept)  Residual
%StdDev:    1.284666 0.8712694
%
%Variance function:
% Structure: fixed weights
% Formula: ~invwt 
%Fixed effects: y ~ 0 + x1 
%      Value Std.Error  DF  t-value p-value
%x1 5.857223  1.008231 140 5.809405       0
%
%Standardized Within-Group Residuals:
%        Min          Q1         Med          Q3         Max 
%-8.55346135  0.07317618  0.17063844  0.35776723  2.15976654 
%
%Number of Observations: 150
%Number of Groups: 10 
%\end{verbatim}
